R Notebook

### 2018-05-14 analysis/final_analysis.Rmd
Add final expt. 2 figs & analyses

1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
3. Change in informativity over time (are people already being informative on first round? are people 'learning' to be more informative over course of expt? or getting less informative because tired of task?)
4. Just length of referring expression in words?

@@

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, location_abusers, confused)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames))
```

## Check basic (no-hidden) pragmatic effect

@@

@@

## Message length collapsing across all subjects

@@

## Scatter to look at within-subj effects

@@

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no))

hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

```{r}

ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1, text = element_text(size=20))

@@

@@

```{r}
summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
             data = d %>% 
               filter(context == 'far') %>%
               filter(!(gameid %in% all_the_things_ers))
               ))
t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
```

@@

```{r}
summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d %>% filter(!(gameid %in% all_the_things_ers))))
```

helmert coding way (one way of slicing)

```{r}
d.regression <- d %>% 
  unite(condition, context, hidden) %>%
  ungroup() %>%
  mutate(condition = ordered(condition, levels = c('far_no', 'close_no', 'far_yes', 'close_yes')))
contrasts(d.regression$condition) <- contr.helmert(4)
summary(lmer(numRawWords ~ condition + (1 + condition | gameid), data = d.regression))
```

@@

Note: Random intercept doesn't converge (less variance in these numbers...)

```{r}
summary(lmer(numFeatures ~ context * occlusions + (1 + context + occlusions | gameid), data = d))
```

```{r}
summary(glmer(colorMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

```{r}
summary(glmer(textureMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

@@

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_manual(values=cbPalette)

>>>>

1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
3. Change in informativity over time (are people already being informative on first round? are people 'learning' to be more informative over course of expt? or getting less informative because tired of task?)

@@

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
nonNative <- unique((exp2_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames)) %>%
  mutate(occlusions = as.factor(occlusions),
        context = as.factor(context))
```

```{r}
paste0('recruited ', length(exp2_subjInfo$gameid))
paste0('total games ', length(unique(d_annotated$gameid)))
paste0('incomplete ', length(incompleteIDs))
paste0('confused ', length(setdiff(confused,incompleteIDs)))
paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
paste0('final sample ', length(unique(d$gameid)))
```

## Quick listener analysis: how common are errors?

Errors rare...

```{r}
exp2_clicks <- read_tsv('../data/final/experiment2/planned_sample/clickedObj/clickedObj.csv') %>%
  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
  filter(!(gameid %in% badGames))

## Errors by condition
tmp <- exp2_clicks %>%
  group_by(context, occlusions) %>%
  summarize(m = sum(correct), total = length(correct) - m)

chisq.test(t(tmp[,3:4]))

## Error distribution by pair
exp2_clicks %>%
  group_by(gameid) %>%
  summarize(numErrors = 24- sum(correct)) %>%
  group_by(numErrors) %>%
  tally()
```

## Visualize basic (no-hidden) pragmatic effect

@@

summary(lmer(numRawWords ~ context + (1 + context | gameid), d %>% filter(hidden == 'no')))

@@

## Visualize message length collapsing across all subjects

@@

## Visualize individual differences

@@

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no)) 
```

```{r}
ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1, text = element_text(size=20))

@@

Look for population who mentioned all the things all the time...

```{r}
hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

@@

```{r}
# summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
#              data = d %>% filter(context == 'far') 
))
#t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
```

@@

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)

summary(lmer(numRawWords ~ context * occlusions + (1 + context * occlusions | gameid), data = d ))
```

@@

Note: Random interaction doesn't converge (less variance in these numbers...)

```{r}
```

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)
summary(glmer(colorMention ~ occlusions * context + (1 + occlusions | gameid), 
              family = 'binomial', data = d))
```

```{r}
summary(glmer(textureMention ~ occlusions * context + (1 + occlusions | gameid), 
              family = 'binomial', data = d))
summary(tmp2)
fixef(tmp2)
baseprob = fixef(tmp2)[1] 
cat('no-occlusion:', exp(baseprob) / (1 + exp(baseprob)))
occludedprob = fixef(tmp2)[1] + fixef(tmp2)[2]
cat('with occlusion:', exp(occludedprob) / (1 + exp(occludedprob)))
```

@@

ggplot(d.booted, aes(x = context, y = mean, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_manual(values=cbPalette)
